Spatiotemporal Data Classification

### 2018-01-06 Report/BigDataHomeworkPandas.tex
Update report with pandas

	\subsection{Question C}
	This task is implemented in the file DataVisualization.py and the created output is stored in the output folder in the gmplots directory. The requirement is to visualize five of the given train trips. For this purporse, after having excluded all trips with null journey\_id and with less than 1 lists with coordinates, the trips list is shuffled and then iterated only for 5 trips. Afterwards, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).

	Finally, the function write\_group\_gml from the utils.py file is called. In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is uded so as to visualize the selected trips. Below, there is an example of a trip shown on map is presented:

@@

	\section{Exercise 2}
	The files implementing the requirements of this part are in the folder SecondQuestion.

@@

	The code for this question is in the file NearestNeighbours.py. Firstly, all test trips in the test file are extracted and stored in the same data structure as the trip list that contains the training dataset. Subsequently, for each test trip all nearest neighbours are calculated in the function find\_nearest\_neighbours\_for\_test\_trip, keeping the time needed in a variable. At the beginning of the trips processing, the tic method from utils.py file is used (finds the current time in millisec using the datetime.datetime.now() method). After having found the nearest neighbours for the given test trip, the method tictoc() is called, returning the required time. In order to find the nearest neighbours, the following process is implemented: For the test trip and all the training trips, the longitudes and latitudes are stored in tuples which are provided to the function calculate\_dynamic\_time\_warping in order to find the distance between two points. The calculated distances are stored in a list, which finally is sorted in ascending order and only the five nearest neighbours are kept.

	Since now the nearest neighbour trips are found, the trips list is updated (function: get\_updated\_trips\_list) containing now only those five trips. The final step now is to visualize the six trips. For this purpose, three list are created: a list with all the points (longitudes and latitudes), a list with the labels and a list with the colors (in this question the only color is the blue). All this data is provided to the function visualize\_paths in the utils.py file in order to export the trips as map images.

@@

	\subsection{Question A2}
	Following the same logic as above, the given test file(test\_set\_a2.csv) is parsed storing the data in the usual format and the function find\_similar\_subroots\_per\_test\_trip is used to determine the nearest sub-routes for each test trip.

	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the utils.py file. To begin with, the method calc\_lcss is called. In this function, the list of lonlat coordinate tuples are given as input. We keep all similar point values and indexes in two lists while iterating the two given lists. The distance between two points is calculated using the method calculate\_lonlat\_distance in the CleanData.py file. Based on the exercise requirements, two points are considered as equals if their distance is less than 200 metres. Therefore, if the points are equal, and the indexes are the first point of one of the given lists the cell is marked to unit cost and a new similar sequence is started. Otherwise, we continue to an existing similar sequence, increasing by 1 its length. \todo{explain z variable} On the other hand, if points are not equal, then a new subsequence should start, and therefore L[i][j] is set to zero.

@@

	Finally, a similar data preprocessing is followed before visualize the data. This time the points are separated so as to show which of them belong to the nearest sub-route and color them with red color. The function for the visualization is the same that was used for the question 2A1.

	\subsection{Question B}
	The purpose of this task was to export features for classification for the training dataset. In short, a grid should be created and the points of each trip trace should be replaced by a specific square of the grid. The output file where the exported features are stored is the tripFeatures.csv and the tripFeatures.pickle, which was used in order to keep the final data structure with the features.

@@

	Afterwards, all points in the trip list should be replaced by the cell names that each point corresponds to. This logic is implemented in the method replace\_points. For each trip in the list, all coordinates are transformed in tuples format and using the find\_index function a name is mapped to this point. After having replaced all points with their new names, this list is exported in .csv and .pickle format. In order to be able to easily check the created grid, the method visualize\_grid is implemented so as to plot the grid.

	\subsection{Question C}
	The file corresponding to this question is the JourneyClassification.py. First of all, the data structure is loaded from the .pickle file, created in above question as we have already mentioned. In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number. Then it is used in order to split our list to the training and test datasets.
	\todo{explain preprocessdata function}

	It worths to be mentioned here, that method asarrays from numpy was used to the necessary for this question lists. In addition, a list with the names of the three classifiers is created and for each of them, the relevant function is called in order to train the classifier and then count the accuracy.

>>>>

	\subsection{Question C}
	This task is implemented in the function visualize\_trips and the created output is stored in the output folder in the gmplots directory. The requirement is to visualize five of the given train trips. For this purporse, the trips list is shuffled in order to visualize random trips and then iterated only for 5 trips. Afterwards, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).

	Finally, the function write\_group\_gml from the utils\_pandas.py file is called. In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is used so as to visualize the selected trips. Below, there is an example of a trip shown on map is presented:

@@

	\section{Exercise 2}
	The files implementing the requirements of this part are in the file question2\_pandas.py and some functions implementing the logic for this question are in the files in the folder question 2.

@@

	In the function question\_a1, two dataframes are created for both the test and the training trips. The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighboursPandas.py is called in order to get the nearest neighbours, calculating also the time needed for this action.

	\todo{paralization}

	In this function, the 3rd column of the dataframe is providing and it is transformed in tuples with the latitudes and longitudes. The list nearest neighbours will keep in tuples for each train trip the trip\_id, the calculated distance (which is calculated using the dynamic time wraping algorithm), the journeyId and finally all the points of the journey. 
	\todo{custom algorithm or library}

	The final list is sorted and only the 5 nearest neighbours are kept for visualization. preprocessing\_for\_visualization, three list are created: a list with all the points (longitudes and latitudes) of the test trip and its neighbours, a list with the labels and a list with the colors (in this question the only color is the blue). All this data is provided to the function visualize\_paths in the utils\_pandas.py file in order to export the trips as map images.

@@

	\subsection{Question A2}
	Following the same logic as above, in the file question2\_pandas and in the function question\_a2, two dataframes are created for the test and the training dataset using pandas library. The test dataframe is iterated row by row, in order to find the 5 trips from the training dataset which have the longest common subsequence with the specific test trip.

	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the utils.py file. To begin with, the method calc\_lcss is called. In this function, the list of lonlat coordinate tuples are given as input. We keep all similar point values and indexes in two lists while iterating the two given lists. The distance between two points is calculated using the method calculate\_lonlat\_distance in the question1\_pandas.py file. Based on the exercise requirements, two points are considered as equals if their distance is less than 200 metres. Therefore, if the points are equal, and the indexes are the first point of one of the given lists the cell is marked to unit cost and a new similar sequence is started. Otherwise, we continue to an existing similar sequence, increasing by 1 its length. \todo{explain z variable} On the other hand, if points are not equal, then a new subsequence should start, and therefore L[i][j] is set to zero.

@@

	Finally, a similar data preprocessing is followed before visualize the data. This time the points are separated so as to show which of them belong to the nearest sub-route and color them with red color. The function for the visualization is the same that was used for the question 2a1 in the utils\_pandas.py file.

	\subsection{Question B}
	The purpose of this task was to export features for classification for the training dataset. In short, a grid should be created and the points of each trip trace should be replaced by a specific square of the grid. The output file where the exported features are stored is the tripFeatures.csv.

@@

	Afterwards, all points in the trip dataframe should be replaced by the cell names that each point corresponds to. This logic is implemented in the method replace\_points. For each row in the dataframe, all coordinates are transformed in tuples format and using the find\_index function a name is mapped to this point. After having replaced all points with their new names, the dataframe is exported in .csv format. In order to be able to easily check the created grid, the method visualize\_grid is implemented so as to plot the grid.

	\subsection{Question C}
	In the method question\_c in question2\_pandas.py file, a new dataframe is created from the features file, containing the grid cell names. All functions used in this question are in the file JourneyClassificationPandas.py. Before starting the classification, the method preprocess\_data was used in order to get the journey\_id of all trips as numbers and also to remove the letter C from the cell names.

	In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number and then the training dataset is splitted in training and test folds. It worths to be mentioned here, that method asarrays from numpy was used to the necessary for this question lists. In addition, a list with the names of the three classifiers is created and for each of them, the relevant function is called in order to train the classifier and then count the accuracy.
